EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-04-04 edipack2.tex
Updated draft up to EDIpack2ineq Going to results, skip other sections

\section{Inequivalent impurities}

\section{C-bindings}\label{SecCbind}

\section{Python API}

\section{Triqs interface}\label{SecEDI2Triqs}

\section{W2Dynamics interface}\label{SecEDI2W2Dyn}

@@

\section{Conclusions}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Inequivalent impurities: EDIpack2Ineq}\label{SecIneq}
It is often necessary to solve systems with several, independent,
quantum impurities. Paradigmatic examples arise in the DMFT
framework when dealing with the effective description of lattices
built out of unit cells hosting multiple inequivalent atoms or large
super-cell systems with possibly broken translational symmetries,
e.g. heterostructures of correlated materials, disordered systems,
etc.
As the structure of \NAME allows the presence of a single instance of
the solver at any time, we implemented a suitable extension to deal
with multiple impurities. Specifically {\tt EDIpack2Ineq} is a
sub-library which complements \NAME, wrapping memory and procedures in
order to enable simultaneous treatment of  inequivalent impurities. 
Using standard Fortran interfaces it is possible to group all the
\NAME and {\tt EDIpack2ineq} functions under the same name without
compromising the software functionalities.  
Here below we briefly discuss the main additions introduced by {\tt
  EDIpack2ineq}.

\subsection{Structure}\label{sSecIneqStructure}
The sub-library is composed of a number of Fortran modules which are
wrapped by {\tt EDIPACK2INEQ}. This latter represents the main user
interface and gives access to all the available procedures and
variables as needed to solve quantum impurity problems.
The user needs to invoke use of this module alongside the \NAME one:
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
program test
  !Load EDIpack2 library 
  USE EDIPACK2 
  !Load the Inequivalent impurities extension
  USE EDIPACK2INEQ
  ...
\end{lstlisting}

A major part of the {\tt EDIpack2ineq} library is to define global
variables which wrap and conserve the memory pool of \NAME.
For instance we implement a number of
globally shared arrays with an increased rank to store the
variables in any memory instance of \NAME and accessible through
the input/output procedures (see \secu{sSecIO}).
Specific objects, e.g. {\tt effective\_bath} or {\tt gfmatrix}, in
principle require dedicated MPI
communication when parallel execution of the solver among independent
impurities. However, we opt for the much simpler and economic solution
of reading these objects from the file in which they get saved
anyway.

\subsection{Core routines}\label{sSecIneqGlobal}
The module {\tt E2I\_MAIN} wraps the extensions to the main algorithms
of \NAME, namely: initialization, diagonalization and finalization,
preserving the original naming convention.  
The initialization process is handled by {\tt ed\_init} which is
extended to accepts a bath array or rank-2 (as opposed to rank-1 array
in the \NAME implementation). The leading dimension now determines the
number of inequivalent impurity problems to be solved. This procedure
initialize all the {\tt EDIpack2ineq} variables and prepare each bath
for the diagonalization step. 
In {\tt ed\_solve}, which accepts the same rank-2 bath array, performs the
diagonalization of each quantum impurity problem in either serial of
parallel mode. The parallel mode is further controlled by the input
function {\tt mpi\_lanc=T/F}. If true Lanczos diagonalization is
performed in parallel and inequivalent sites are treated
sequentially. The inverse if false.   
The finalization procedure with full memory release is performed in
{\tt ed\_finalize}.

\subsection{Inequivalent Baths}\label{sSecIneqBath}
The inequivalent impurities extension is encoded in {\tt E2I\_BATH}
which contains procedures to control the baths setup on the user side
and the  implementation of conventional symmetry operations.
In the {\tt E2I\_BATH\_REPLICA} module we extend all the functions
which set the replica/general bath matrix basis 
$\vec{\Gamma}$ and the corresponding initial variational parameters
$\vec{\lambda}$. Although the variational parameters are generally
different from site to site, in the current implementation the matrix
basis is independent from the impurities number. This will possibly be
relaxed in a future release. 
Analogously, in the module {\tt E2I\_BATH\_FIT} we implement an extension of the generic
function {\tt ed\_chi2\_fitgf} to perform multiple independent
bath optimization for all inequivalent impurities in terms of an MPI iterative cycle.

\subsection{Input/Output}\label{sSecIneqIO}
A key part of the {\tt EDIpack2ineq} library is the extension for the
input/output procedures to return observables or dynamical functions per inequivalent impurity.
The module {\tt E2I\_IO} implements numerous functions which retrieve
or evaluate on-the-fly any quantity accessible in \NAME for any given
impurity (or all of them) using the same naming original convention
defined in \NAME.

\section{C-bindings: EDIpack2C}\label{SecCbind}

\section{Python API: EDIpy2}\label{SecPyAPI}

\section{Triqs interface: EDIpack2Triqs}\label{SecEDI2Triqs}

\section{W2Dynamics interface: EDIpack2w2dyn}\label{SecEDI2W2Dyn}

@@

\section{Conclusions and perspectives}
\subsection{Conclusions}

\subsection{Perspectives}
